Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2023-11-27 article.qmd
Add text for the reproducibility and runtime sections

![Performance of the different softwares to solve a subset of BioModels models. (left) Run time for model BIOMD3 as a function of the number of output points. (right) Run time for different models ordered according to the geometric mean of the run times.](src/performance/figures/performance.png){#fig-runtime}

>>>>

![Performance of the different softwares to solve a subset of BioModels models. (left) Run time for model BIOMD3 as a function of the number of output points. (right) Run time for different models ordered according to the geometric mean of the run times, using 300 output points.](src/performance/figures/performance.png){#fig-runtime}

As poincaré and SimBio are coded in pure Python,
an expected concern is its runtime performance when integrating a model.
To test it,
we downloaded 28 models from BioModels,
and simulated them with COPASI, Tellurium/RoadRunner and simbio.
For simbio, we considered both the pure Python and numba-compiled code.
We used the LSODA solver with $10^-6$ relative and absolute tolerances,
with the exception of tellurium,
which provides the similar CVODE solver.
We measured separately the loading, first/cold run and subsequent/warm runs of each model.

For COPASI and Tellurium/RoadRunner,
we noted that its runtime depended on the number of evaluation points,
something that does not happen with SimBio (@fig-runtime, left).
Using numba to compile the right-hand-side (RHS) code generated by SimBio
results in an order of magnitude speed-up,
putting its performance on par with COPASI and RoadRunner.
If we switch the LSODA `scipy` solver for a more efficient `numbalsoda` implementation,
we get another speed-up in the SimBio runtime,
beating all other methods.
This is also true for the other models we tested (@fig-runtime, right).
